2.2 Rogen-Gladen Adjustment

### 2023-12-18 2_scripts/2.2 Rogen-Gladen Adj.qmd
add anti-S figures

## Load necc paks
```{r}
library(dplyr)         # Data management
library(reshape2)      # for melt() function
library(ggplot)        # ggplot
library(zoo)           # year-month format
```
# ============= Part I. The Changing of Anti-N Pos APL ===============#

@@

>>>>

## Load necc paks
```{r}
library(dplyr)         # Data management
library(reshape2)      # for melt() function
library(ggplot2)        # ggplot
library(zoo)           # year-month format
```
# ============= Part I. The Changing of Anti-N Pos APL ===============#

@@

##==================== Part III. Anti-S Positivity Overtime ================##

#### This time we will use anti-S only from the CBS data, since among-data diff is not the major focus of this project

```{r}
# Roche_s: like other Roche assay, COI = 0.4, those < 0.4 = negative and recorded as 0.4

# Histogram 
hist(CBS_Ab$roche_s[CBS_Ab$roche_s>0.4]) # extreme values
summary(CBS_Ab$roche_s) # This time they have 10,0000 as max, much larger than CLSA's 2500

# Histogram log-transformed
hist(CBS_Ab$log_roche_s[CBS_Ab$roche_s>0.4]) # Aman did log_e transform
summary(CBS_Ab$log_roche_s)

# Binary
table(CBS_Ab$interp_roche_s, useNA = 'ifany')
# Majority (87.50%) were test positive
```
#### Calculate Anti-S Positivity
```{r}
# prepare melt data
tb1<-CBS_Ab%>%
  group_by(month_year) %>%
  summarise(n = n(), )

tb2<-CBS_Ab[CBS_Ab$interp_roche_s==1,] %>%
  group_by(month_year) %>%
  summarise(s.pos = n(), )

tb1<-merge(tb1, tb2, by = "month_year", all.x = T, all.y = F)
tb1$prop.s<-tb1$s.pos/tb1$n
tb1$month_year<-as.yearmon(tb1$month_year)
tb1<-tb1[order(tb1$month_year),]

# Third-party sens & spec
# Sen=97.95, spec = 99.95 
# Source:https://www.sciencedirect.com/science/article/pii/S016609342100210X
tb1$prop.s_RG_l<- (tb1$prop.s+1-0.9995)/(1+0.9795-0.9995)

CBS_melt2<-melt(select(tb1, month_year, prop.s, prop.s_RG_l), id.vars = "month_year")

p02<-CBS_melt2 %>%
  ggplot(aes(x=month_year, y=value, colour=variable)) +
  geom_point(aes(colour=variable)) +
  geom_line(aes(colour=variable))  +
  geom_hline(yintercept=0, colour='red', linetype=2) +
  xlab("Time") + ylab("Estimated Prevalence") +
  theme_bw()

p02$labels$colour<-"Adjustment"
p02
```
#### Add additional COI 0.8 and 15

```{r}
CBS_Ab$interp_roche_s0.8<-if_else(CBS_Ab$roche_s>0.8, 1, 0)

# elevating COI from 0.4 to 0.8 only increase N(-) by 2

CBS_Ab$interp_roche_s15<-if_else(CBS_Ab$roche_s>15, 1, 0)
table(CBS_Ab$interp_roche_s15, useNA = 'ifany')

# elevating COI from 0.4 to 15 increase N(-) by 2000
```
#### Plot 03:  0.4 and 15 COI together in one single plot
```{r}
# Creating the N_prop data
tb1<-CBS_Ab%>%
  group_by(month_year) %>%
  summarise(n = n(), )

tb2<-CBS_Ab[CBS_Ab$interp_roche_s==1,] %>%
  group_by(month_year) %>%
  summarise(s.pos = n(), )

tb3<-CBS_Ab[CBS_Ab$interp_roche_s15==1,] %>%
  group_by(month_year) %>%
  summarise(s.pos15 = n(), )

tb1<-merge(tb1, tb2, by = "month_year", all.x = T, all.y = F)
tb1<-merge(tb1, tb3, by = "month_year", all.x = T, all.y = F)

tb1$prop.s<-tb1$s.pos/tb1$n
tb1$prop.s15<-tb1$s.pos15/tb1$n
tb1$month_year<-as.yearmon(tb1$month_year)
tb1<-tb1[order(tb1$month_year),]
```

```{r}
# Melt the data
CBS_melt3<-melt(select(tb1, month_year, prop.s, prop.s15), id.vars = "month_year")

p03<-CBS_melt3 %>%
  ggplot(aes(x=month_year, y=value, colour=variable)) +
  geom_point(aes(colour=variable)) +
  geom_line(aes(colour=variable))  +
  geom_hline(yintercept=0, colour='red', linetype=2) +
  xlab("Time") + ylab("Estimated Prevalence") +
  theme_bw()

p03$labels$colour<-"COI"
p03
```
#### Add RG-adjustment to the plot, using the sens and spec related to detection of virus neutrolization, rather than detection of presence of Antibodies.
```{r}
# N and proportions
tb1<-CBS_Ab%>%
  group_by(month_year) %>%
  summarise(n = n(), )

tb2<-CBS_Ab[CBS_Ab$interp_roche_s0.8==1,] %>%
  group_by(month_year) %>%
  summarise(s.pos0.8 = n(), )

tb3<-CBS_Ab[CBS_Ab$interp_roche_s15==1,] %>%
  group_by(month_year) %>%
  summarise(s.pos15 = n(), )

tb1<-merge(tb1, tb2, by = "month_year", all.x = T, all.y = F)
tb1<-merge(tb1, tb3, by = "month_year", all.x = T, all.y = F)

tb1$prop.s0.8<-tb1$s.pos0.8/tb1$n
tb1$prop.s15<-tb1$s.pos15/tb1$n
tb1$month_year<-as.yearmon(tb1$month_year)
tb1<-tb1[order(tb1$month_year),]
```

```{r}
# Add R-G adjusted props
# Source: https://diagnostics.roche.com/global/en/products/params/elecsys-anti-sars-cov-2-s.html
# Sample size is small, find a larger one?

tb1$prop.s0.8_RG<- (tb1$prop.s0.8+0.875-1)/(0.875+0.994-1)
tb1$prop.s15_RG<- (tb1$prop.s15+1-1)/(1+0.925-1)

# melt the data
CBS_melt4<-melt(select(tb1, month_year, prop.s0.8, prop.s15,
                       prop.s0.8_RG, prop.s15_RG), id.vars = "month_year")
CBS_melt4$adj[CBS_melt4$variable %in% c("prop.s0.8", "prop.s15")]<-"no adjustment"
CBS_melt4$adj[CBS_melt4$variable %in% c("prop.s0.8_RG", "prop.s15_RG")]<-"RG adjusted"
table(CBS_melt4$variable, CBS_melt4$adj)

CBS_melt4$COI[CBS_melt4$variable %in% c("prop.s0.8", "prop.s0.8_RG")]<-"0.8"
CBS_melt4$COI[CBS_melt4$variable %in% c("prop.s15", "prop.s15_RG")]<-"15"
table(CBS_melt4$variable, CBS_melt4$COI)
```
```{r}
## ggplot using melted data
p04<-CBS_melt4 %>%
  ggplot(aes(x=month_year, y=value, colour=adj)) +
  geom_point(aes(colour=adj, shape=COI)) +
  geom_line(aes(colour=adj, linetype=COI))  +
  geom_hline(yintercept=0, colour='red', linetype=2) +
  xlab("Time") + ylab("Estimated Prevalence") +
  theme_bw()

p04$labels$colour<-"Adjustment"

p04

# plot COI = 0.8/15 seperately
```
